The complex network of trophic interactions in a subAntarctic oceanic Marine Protected Area

### 2024-03-27 MS_ToSubmit/RevisedVersion_2.Rmd
Abstract length

The world’s oceans designated under marine protection have increased recently. Most Marine Protected Areas (MPAs) target vulnerable, keystone, charismatic and or endemic species. In the sub-Antarctic, ocean protection is associated to oceanic islands, except for MPAs Namuncurá - Burdwood Bank I and II (MPA N-BB, ~53º–55ºS and ~56º–62ºW), which are associated to a submarine plateau and a southern deep slope, respectively. Here, we present the first analysis of the predator-prey network for the MPA N-BB. We applied a topological network approach to characterise the complexity and structure of the food web, and identify the species’ role in such a framework. The MPA N-BB food web consisted of 1788 interactions and 379 species, with a connectance of 0.01. Almost half of the consumers feed at more than one trophic level (0.48), and the network displayed a small-world pattern (short path length, high clustering of compartments). This network pattern suggests that the ecosystem might be vulnerable to perturbations targeting highly connected species, although some properties might provide resilience and resistance, resulting in a rearranged structure that preserves its original functions. Several species arose as important for the trophic structure and functioning, and response to perturbations. Generalist species, mainly fishes, play a crucial role in the benthopelagic coupling and should be considered as relevant energy transfers for the ecosystem. We argue that the diversity of species, including both the benthic and pelagic habitats, is responsible for securing the connectivity within the food web against perturbations, therefore contributing to the structure and stability of the ecosystem.

>>>>

The world’s oceans designated under marine protection have increased recently. Most Marine Protected Areas (MPAs) target vulnerable, keystone, charismatic and/or endemic species. In the sub-Antarctic, ocean protection is associated to oceanic islands, except for MPAs Namuncurá - Burdwood Bank I and II (MPA N-BB, ~53º – 55ºS and ~56º – 62ºW), which are associated to a submarine plateau and a southern deep slope, respectively. We present the first analysis of the predator-prey network for the MPA N-BB, applying a topological network approach to characterise the complexity and structure of the food web, and identify the species’ role. The MPA N-BB food web consisted of 1788 interactions and 379 species, with a connectance of 0.01. Almost half of the consumers feed at more than one trophic level (0.48), and the network displayed a small-world pattern (short path length, high clustering of compartments). This network pattern suggests that the ecosystem might be vulnerable to perturbations targeting highly connected species, although some properties might provide resilience and resistance, resulting in a rearranged structure that preserves its original functions. Several species arose as important for the trophic structure and functioning, and response to perturbations. Generalist species, mainly fishes, play a crucial role in the benthopelagic coupling and should be considered as relevant energy transfers for the ecosystem. We argue that the diversity of species, including both the benthic and pelagic habitats, is responsible for securing the connectivity within the food web against perturbations, therefore contributing to the structure and stability of the ecosystem.
